---
title: "Parse LECS Files"
format: html
editor: source
---

Get and parse data from LECS files.



# Get Data

```{r}
#| warning: false
#| message: false

library(data.table)
library(tidyverse)
library(mlabtools)
library(clock)
library(tictoc)
library(furrr)
library(parallel)

theme_set(theme_bw())
options(digits.secs=3)

# RINKO_CALS
rinko_cals <- list(
  temp_A = -1.219367e1,
  temp_B = 2.134089e1,
  temp_C = -3.559172e00,
  temp_D = 6.691104e-01,
  o2_A = -4.382235e01,
  o2_B = 1.398755e02,
  o2_C = -4.119456e-01,
  o2_D = 9.934000e-03,
  o2_E = 4.000000e-03,
  o2_F = 4.440000e-05,
  o2_G = 0.000000e+00,
  o2_H = 1.000000e+00
)

# plotting function
plot_mean <- function(data, parameter, err) {
  data |> 
    ggplot(aes(x = timestamp_mean, 
               y = {{parameter}},
               ymin = {{parameter}} - {{err}},
               ymax = {{parameter}} + {{err}})) +
    geom_line(color = "lightgray") +
    geom_linerange(color = "gray") +
    geom_point(size = 1) +
    labs(x = NULL)
}
```

## Load data

```{r}
met <- fread("data/processed/surface/lecs_met_2023-09-05.csv")
status <- fread("data/processed/surface/lecs_status_2023-09-05.csv")
adv_data <- fread("data/processed/surface/lecs_adv_data_2023-09-05.csv")
```

Remove lines with duplicate timestamps

how many duplicates?

```{r}
length(unique(adv_data$timestamp)) / nrow(adv_data)
```

remove duplicates

```{r}
adv_data <- distinct(adv_data, timestamp, .keep_all = TRUE)
fwrite(adv_data, "data/processed/surface/lecs_adv_data_2023-09-05d.csv")
```


Add missing counts between lines to ADV data

```{r}
adv_data <- adv_data |> 
  mutate(missing = lecs_missing(count))
```

Add timestamp from nearest status line to ADV data. Doesn't work with lots of data.

```{r}
#| eval: false
find_nearest_row <- function(row_number) {
  return(which.min(abs(status$row_num - row_number)))
}

adv_data$timestamp_s <- status$timestamp[sapply(adv_data$row_num, find_nearest_row)]
```

Filter bad data

```{r}
status_qc <- status |> 
  filter(soundspeed > 1450, 
         adv_day < 32, adv_month > 0, adv_month < 13, 
         adv_min < 61, adv_hour < 24, adv_year < 100,
         timestamp > "2023-01-01",
         timestamp < "2024-10-01")


adv_data_qc <- adv_data |> 
  filter(count >= 0, count < 256, 
         ana_in2 == 1, 
         timestamp > "2023-01-01",
         timestamp < "2024-10-01")

```

Calculate per-hour means

```{r}
met_mean <- met |> 
  group_by(date_group(timestamp, "hour")) |> 
  summarise(across(everything(), 
                   list(mean = ~ mean(.x, na.rm = TRUE), 
                        sd = ~ sd(.x, na.rm = TRUE))))

status_mean <- status_qc |> 
  select(timestamp, bat) |> 
  group_by(date_group(timestamp, "hour")) |> 
  summarise(across(c(timestamp, bat), 
                   list(mean = ~ mean(.x, na.rm = TRUE), 
                        sd = ~ sd(.x, na.rm = TRUE))))

adv_data_mean <- adv_data_qc |> 
  select(missing, timestamp, pressure, pH, temp, oxy) |> 
  group_by(date_group(timestamp, "hour")) |> 
  summarise(across(everything(), 
                   list(mean = ~ mean(.x, na.rm = TRUE), 
                        sd = ~ sd(.x, na.rm = TRUE))),
            missing_frac = sum(missing, na.rm = TRUE) / 
                           (sum(missing, na.rm = TRUE) + n()),
            N = n(),
            missing_frac_t = 1 - N/(60*60*16))

```

### Met data

#### PAR

```{r}
met_mean |> 
  plot_mean(PAR_mean, PAR_sd)
```

#### Wind Speed

```{r}
met_mean |> 
  plot_mean(wind_speed_mean, wind_speed_sd) +
  labs(title = "Wind Speed",
       y = "Wind Speed (m/s)")
```

#### Wind direction

```{r}
ggplot(met_mean, aes(wind_dir_mean, wind_speed_mean)) +
  geom_point() +
  coord_polar() 
```

### Status

#### Battery

```{r}
status_mean |> 
  plot_mean(bat_mean, bat_sd) +
  geom_smooth() +
  labs(title = "Battery Voltage",
       y = "Volts")
```

### ADV Data

#### Check time alignment

```{r}
#ggplot(adv_data, aes(row_num, timestamp)) + geom_point()
```

Compare expected time diffs. Comparing diff from counts to timestamp diff should be 0. Diffs indicate something unexpected happening with time alignment.

```{r}
adv_data <- adv_data |> 
  mutate(time_diff_missing = (missing + 1)/16,
         time_diff_timestamp = timestamp - lag(timestamp),
         time_diff_diff = time_diff_missing - time_diff_timestamp)
```

```{r}
adv_data |> 
  head(5000) |> 
  filter(abs(time_diff_diff) < 1) |> 
  ggplot(aes(timestamp, time_diff_diff)) + 
  geom_point()
```


#### pressure

```{r}
adv_data_mean |> 
  plot_mean(pressure_mean, pressure_sd) +
  labs(title = "Pressure",
       y = "Pressure (dbar)")
```

#### pH

Data issues cause wild flyers

```{r}
#| eval: false
adv_data_qc |> 
  ggplot(aes(time, pH)) +
  geom_point()
```


```{r}
adv_data_mean |> 
  plot_mean(ph_counts_mean, ph_counts_sd) +
  labs(title = "pH")
```

#### Temperature

```{r}
adv_data_mean |> 
  plot_mean(temp_mean, temp_sd) +
  labs(title = "Temperature",
       y = "Temp (C)")
```

#### Oxygen

```{r}
adv_data_mean |> 
  plot_mean(DO_percent_mean, DO_percent_sd) +
  labs(title = "Oxygen")
```

## Missing data

Missing data by row. Determined by looking at gaps in cycle number.

```{r}
#| eval: false
adv_data |> 
  ggplot(aes(row_num, missing)) +
  geom_point() +
  labs(title = "Missing data by row",
       x = NULL,
       y = "Missing")
```

Fraction of data missing. Ratio of total missing rows (as above)
to expected rows (rows in send + sum of missing rows).

```{r}
adv_data_mean |> 
  ggplot(aes(timestamp_mean, missing_frac)) +
  geom_point() +
  labs(title = "Fraction missing by minute",
       x = NULL,
       y = "Fraction missing")
```

The same calculated as actual number of rows / expected rows (60s * 16rps)

```{r}
adv_data_mean |> 
  ggplot(aes(timestamp_mean, missing_frac_t)) +
  geom_point() +
  labs(title = "Fraction missing by minute",
       x = NULL,
       y = "Fraction missing")
```