---
title: "2023 LECS data"
format: html
---

Exploratory plots of 2023 LECS data

## load libraries and read data

```{r}
library(tidyverse)
library(mlabtools)
library(scattermore)
library(xts)
library(dygraphs)

met <- read_csv("data/LECS_2023/surface_met.csv")
sfc_status <- read_csv("data/LECS_2023/surface_status.csv")
sfc_data <- read_csv("data/LECS_2023/surface_data.csv")
ldr_status <- read_csv("data/LECS_2023/lander_status.csv")
ldr_data <- read_csv("data/LECS_2023/lander_data.csv")
```

ADV had incorrect RTC setting. Adding an offset to convert to clock time

```{r}
# time offset in seconds
ts_offset <- 671628945

adv_status <- read_adv_status("data/LECS_2023/LECS_ADV/22.sen", ts_offset)
```

# ADV data

Data stored in internal ADV memory

## Battery Voltage

Resampled to 1 min means

```{r}
bat_ts <- xts(adv_status$bat, order.by = adv_status$timestamp)
bat_ts_res <- period.apply(bat_ts, endpoints(bat_ts, "mins"), FUN = mean)
bat_ts_res <- bat_ts_res[bat_ts_res[,1] < 19,]
dygraph(bat_ts_res) |> 
  dygraphs::dyRangeSelector()
```

## Temperature

Resampled to 1 min means

```{r}
temp_ts <- xts(adv_status$temp, order.by = adv_status$timestamp)
temp_ts_res <- period.apply(temp_ts, endpoints(temp_ts, "mins"), FUN = mean)
dygraph(temp_ts_res) |> 
  dygraphs::dyRangeSelector()
```


# Met data

## PAR

Something's up with the PAR readings. Is this something electrical related to the 4h transmit cycles?

```{r}
plot_ts(met, "par", timestamp = "ts")
```

### Wind Speed

Wind speed and direction look OK, if a little noisy

```{r}
plot_ts(met, "wind_speed", timestamp = "ts")
```

### Wind Dir

```{r}
plot_ts(met, "wind_dir", timestamp = "ts")
```

# Lander SD Data

Looking at single most recent 1h file.

Missing packets

Estimate missing packet fraction by looking at 0-255 count. Estimated 4% with python.

```{r}
ldr_data_missing <- ldr_data |> 
  filter(count != 0) |> 
  mutate(missed = ifelse(count == 1, 256 - lag(count), count - 1 - lag(count)))
```

Fraction missing

```{r}
sum(ldr_data_missing$missed, na.rm = TRUE) / length(ldr_data_missing$missed)
```

pH

```{r}
ggplot(ldr_data, aes(1:nrow(ldr_data), pH)) +
  geom_line()
```

Estimate bad packet fraction with pH

```{r}
ldr_data_bad <- ldr_data |> 
  mutate(bad = (pH > 8.4 | pH < 7.5))

sum(ldr_data_bad$bad) / length(ldr_data_bad$bad)
```

pH without bad packets

```{r}
df <- ldr_data_bad |> 
  filter(!bad)

ggplot(df, aes(1:nrow(df), pH)) +
  geom_line()
```

DO & temp

These look OK.

```{r}
ggplot(ldr_data, aes(1:nrow(ldr_data), DO)) +
  geom_line()
```

```{r}
ggplot(ldr_data, aes(1:nrow(ldr_data), temp)) +
  geom_line()
```

# Surface SD Status

Data is very noisy. Pretty sure this is due to bad ADV packet reads. 

```{r}
bat_ts <- xts(sfc_status$bat, order.by = sfc_status$ts)
bat_ts_res <- period.apply(bat_ts, endpoints(bat_ts, "mins"), FUN = mean)
bat_ts_res <- bat_ts_res[bat_ts_res[,1] < 13 & bat_ts_res[,1] > 11,] # filter by voltage
bat_ts_res <- bat_ts_res['2023-03-01/2023-11-01'] # filter by time
dygraph(bat_ts_res) |> 
  dygraphs::dyRangeSelector()
```

