---
title: "PAR"
format: html
---

LECS PAR sensor data.

LECS has a surface PAR sensor (LICOR) on the surface buoy
and a submerged sensor on the lander (PAR Odyssey).
Underwater sensor is a logger and should produce
calibrated (?) PAR. 
The LICOR sensor is in units of counts which represent voltage.
It needs calibration, either from the factory data sheet or
by cross-calibration with another sensor.

Matt has a paper on calibrating these and HOBO light sensors:
DOI 10.4319/lom.2012.10.416


```{r}
library(tidyverse)
library(mlabtools)
library(dygraphs)
```

# Underwater PAR

```{r}
lecs_read_par_odyssey <- function(file) {
  header <- c("scan", "date", "time", "raw_par", "cal_par")
  read_csv(file, skip = 9, col_names = header) |> 
    mutate(date = as.Date(date, "%d/%m/%Y"),
           timestamp = lubridate::parse_date_time(paste(date, time), 
                                                  "Ymd HMS",
                                                  tz = "UTC")) |> 
    select(scan, timestamp, raw_par, cal_par)
}
files <- c("data/PAR/LECS_10-5-23.CSV", "data/PAR/LECS_PAR_2024-02-28.CSV")
df <- lecs_read_par_odyssey(files)

ggplot(df, aes(timestamp, cal_par)) + geom_line()
dygraph(select(df, timestamp, cal_par))
```


