---
title: "PAR"
format: html
---

LECS PAR and light sensor data.

LECS has a surface PAR sensor (LICOR) on the surface buoy
and a submerged sensor on the lander (PAR Odyssey).
Underwater sensor is a logger and should produce
calibrated (?) PAR. 
The LICOR sensor is in units of counts which represent voltage.
It needs calibration, either from the factory data sheet or
by cross-calibration with another sensor.

Matt has a paper on calibrating these and HOBO light sensors:
DOI 10.4319/lom.2012.10.416


```{r}
library(tidyverse)
library(mlabtools)
library(dygraphs)
library(arrow)
theme_set(theme_bw())
```

# surface PAR

surface PAR is currently uncalibrated in ADC counts.
Needs calibration in environmental chamber or via codeployed sensor.

Data also needs QC. Some periods need outlier elimination, 
others look unrecoverable.

```{r}
sfc_par <- open_dataset("data/processed/met_all.parquet")
names(sfc_par)
sfc_par <- sfc_par |> 
  select(timestamp, PAR) |> 
  collect()
dygraph(sfc_par)
```

# Underwater PAR

```{r}
files <- c("data/PAR/LECS_10-5-23.CSV", "data/PAR/LECS_PAR_2024-02-28.CSV")
df <- lecs_read_par_odyssey(files)

ggplot(df, aes(timestamp, cal_par)) + geom_line()
```


```{r}
dygraph(select(df, timestamp, cal_par))
```


