---
title: "Resample LECS ADV timeseries"
format: html
editor: source
---

Resample filtered and calibrated data to hourly and every minute

```{r}
library(tidyverse)
library(arrow)
library(dygraphs)
library(tictoc)

options(digits.secs = 6)

data_dir <- "data/processed/"
```

```{r}
ds <- open_dataset(paste0(data_dir, "adv_all_filt_cal.parquet"))
names(ds)
```

resample to hourly

```{r}
ds_hourly <- ds |> 
  select(timestamp, pressure, u, v, w, amp1, amp2, amp3, corr1, corr2, corr3, temp, pH, DO_percent) |>
  mutate(timestamp = floor_date(timestamp, "hour")) |> 
  group_by(timestamp) |> 
  summarise(across(everything(), 
                   ~ mean(.x, na.rm = TRUE))) |> 
  ungroup()

ds_hourly |> 
  collect() |> 
  data.table::fwrite(paste0(data_dir, "adv_all_filt_cal_hourly.csv"))

write_dataset(ds_hourly, 
              paste0(data_dir, "adv_all_filt_cal_hourly.parquet"))
```

resample to minute

```{r}
ds_minute <- ds |> 
  select(timestamp, pressure, u, v, w, amp1, amp2, amp3, corr1, corr2, corr3, temp, pH, DO_percent) |>
  mutate(timestamp = floor_date(timestamp, "minute")) |> 
  group_by(timestamp) |> 
  summarise(across(everything(), 
                   ~ mean(.x, na.rm = TRUE))) |> 
  ungroup()

ds_minute |> 
  collect() |> 
  data.table::fwrite(paste0(data_dir, "adv_all_filt_cal_minute.csv"))

write_dataset(ds_minute, 
              paste0(data_dir, "adv_all_filt_cal_minute.parquet"))
```

```{r}
df_minute <- ds_minute |> 
  collect()
```

```{r}
df_minute |> 
  ggplot(aes(timestamp, pH)) +
  geom_line()
```

Resample to 4Hz and impute missing data

Up to 4 missing samples?

```{r}
ds_4hz <- ds |> 
  select(timestamp, pressure, u, v, w, amp1, amp2, amp3, corr1, corr2, corr3, temp, pH, DO_percent) |>
  mutate(timestamp = floor_date(timestamp, ".25 sec")) |> 
  group_by(timestamp) |> 
  summarise(across(everything(), 
                   ~ mean(.x, na.rm = TRUE))) |> 
  ungroup()

```

create complete 4Hz timeseries and join to add NA's for missing data

```{r}
names(ds_4hz)
df_4hz <- ds_4hz |> 
  arrange(timestamp) |> 
  collect()

head(df_4hz)
tail(df_4hz)

# ts <- collect(ds_4hz$timestamp) # doesn't work
st <- min(df_4hz$timestamp, na.rm = TRUE) 
et <- max(df_4hz$timestamp, na.rm = TRUE) 
ts_4hz <- seq(from = st, to = et, by = 0.25) 
head(ts_4hz)

df_ts_4hz <- tibble(timestamp = ts_4hz)

df_4hz_all <- df_ts_4hz |> 
  left_join(df_4hz)

write_dataset(df_4hz_all, paste0(data_dir, "adv_4hz.parquet"))

head(df_4hz_all, n = 1000) |> View()
```

check missingness with imputeTS

```{r}
library(imputeTS)
```


```{r}
# pick a small "good" subset
pts <- df_4hz_all |> 
  filter(timestamp > "2024-01-01 00:00:00",
         timestamp < "2024-01-01 01:00:00") |> 
  collect() |> 
  pull(pH) |> 
  ts(frequency = 4)

ggplot_na_distribution2(pts)
```

same for full frequency dataset

```{r}
# pick a small "good" subset
df <- ds |> 
  filter(timestamp > "2024-01-01 00:00:00",
         timestamp < "2024-01-01 01:00:00") |> 
  collect()

length(unique(df$timestamp))
```

Uh oh, we have duplicates

```{r}
ggplot(df, aes(1:nrow(df), timestamp)) +
  geom_point()
```

wth? time goes back 30 min for a bit?

how much does this happen over the course of a day?

```{r}
df <- ds |> 
  filter(timestamp > "2024-01-01 00:00:00",
         timestamp < "2024-01-02 00:00:00") |> 
  collect()

length(unique(df$timestamp))
```

Uh oh, we have duplicates again

```{r}
df |> 
  arrange(timestamp) |> 
ggplot(aes(1:nrow(df), timestamp)) +
  geom_point()

```

```{r}
df |> 
  select(timestamp) |> 
  arrange(timestamp) |> 
  mutate(dup = duplicated(timestamp)) |> 
  head(150)
```


```{r}
st <- min(df$timestamp) 
et <- max(df$timestamp) 
df_ts <- tibble(timestamp = seq(from = st, to = et, by = 0.0625))
  
df_ts <- df_ts |> 
  left_join(df)

pts <- df_ts |> 
  pull(pH) |> 
  ts(frequency = 16)

ggplot_na_distribution2(pts)
```

