---
title: "Resample LECS ADV timeseries"
format: html
editor: source
---

* Resample filtered and calibrated data to hourly and every minute
* Resample to 4Hz

impute missing points. Acceptable number?

How to divide up remaining points and group

* group between gaps
* group these groups by 1h or 15 min
* run flux on each group

```{r}
library(tidyverse)
library(arrow)
# library(tictoc)
library(imputeTS)
library(polars)

options(digits.secs = 6)

data_dir <- "data/processed/surface"
```

```{r}
ds <- open_dataset(file.path(data_dir, "lecs_adv_cal.parquet"))
```

resample to hourly

```{r}
ds_hourly <- ds |> 
  select(-c(year, month)) |> 
  mutate(timestamp = floor_date(timestamp, "hour")) |> 
  group_by(timestamp) |> 
  summarise(across(everything(), 
                   ~ mean(.x, na.rm = TRUE))) |> 
  ungroup()

ds_hourly |> 
  collect() |> 
  data.table::fwrite(file.path(data_dir, "lecs_adv_hourly.csv"))

write_dataset(ds_hourly, 
              file.path(data_dir, "lecs_adv_hourly.parquet"))
```

resample to minute

```{r}
ds_minute <- ds |> 
  select(-c(year, month)) |> 
  mutate(timestamp = floor_date(timestamp, "minute")) |> 
  group_by(timestamp) |> 
  summarise(across(everything(), 
                   ~ mean(.x, na.rm = TRUE))) |> 
  ungroup()

ds_minute |> 
  collect() |> 
  data.table::fwrite(file.path(data_dir, "lecs_adv_minute.csv"))

write_dataset(ds_minute, 
              file.path(data_dir, "lecs_adv_minute.parquet"))
```

Resample to 4Hz and impute missing data

create complete 4Hz timeseries and join to add NA's for missing data

```{r}
df_4hz <- ds |> 
  select(-c(year, month)) |> 
  mutate(timestamp = floor_date(timestamp, ".25 sec")) |> 
  group_by(timestamp) |> 
  summarise(across(everything(), 
                   ~ mean(.x, na.rm = TRUE))) |> 
  ungroup() |> 
  arrange(timestamp) |>
  collect()

st <- min(df_4hz$timestamp) 
et <- max(df_4hz$timestamp) 

ts_4hz <- seq(from = st, to = et, by = 0.25) 

df_ts_4hz <- tibble(timestamp = ts_4hz)

df_4hz_all <- df_ts_4hz |> 
  left_join(df_4hz)
```

```{r}
df_4hz_all %>% 
  group_by(year = year(timestamp), month = month(timestamp)) %>% 
  write_dataset(file.path(data_dir, "lecs_adv_4hz.parquet"))
```

Impute by interpolation with a max gap of 1sec.

```{r}
df_4hz_all_imp <- df_4hz_all %>% 
  mutate(across(-timestamp, \(x) na_interpolation(x, maxgap = 4)))

df_4hz_all_imp %>% 
  group_by(year = year(timestamp), month = month(timestamp)) %>% 
  write_dataset(file.path(data_dir, "lecs_adv_4hz_imp.parquet"))
```

Add groups representing continuous data
remove NA's and write to parquet

```{r}
# df_4hz_all_imp <-  open_dataset(file.path(data_dir, "lecs_adv_4hz_imp.parquet")) |> 
#   arrange(timestamp) |> 
#   collect()

df_4hz_imp_grp <- df_4hz_all_imp |> 
  mutate(is_gap = is.na(pH_cal) & !is.na(lag(pH_cal, default = NA))) |> 
  mutate(group = cumsum(is_gap) + 1) |> 
  select(-is_gap) |> 
  drop_na(pressure)
```

```{r}
df_4hz_imp_grp |> 
  group_by(year = year(timestamp), month = month(timestamp)) |> 
  write_dataset(file.path(data_dir, "lecs_adv_4hz_imp_grp.parquet"))
```

divide up big groups

10 min blocks within groups

```{r}
df_grp <- df_4hz_imp_grp  |> 
  group_by(group) |> 
  mutate(block_time = floor_date(timestamp, unit = "10min"),
         block_start = block_time != lag(block_time),
         block_start = if_else(is.na(block_start), TRUE, block_start)) |> 
  ungroup() |> 
  mutate(block = cumsum(block_start) + 1) |> 
  select(-c(block_time, block_start))

```

```{r}
df_grp |> 
  group_by(year, month) |> 
  write_dataset(file.path(data_dir, "lecs_adv_4hz_imp_grp_10.parquet"))
```
